Alignment of Light and Mass in Lensing Galaxies

### 2013-07-06 shapes/paper.tex
paper update

As mentioned in Section \ref{sec:data}, there are different time delay estimates for \textit{PG1115}. We find that the results do not change much when using different estimates. As can be seen in Figure \ref{fig:1115rec} resp. Figure \ref{fig:barkana}, the dark matter distribution is a bit more spherical with the time delays from \cite{2010MNRAS.406.2764T} and the uncertainty on the angle of the semi-major axis grows. The misalignment we see is thus not affected by different time delays.

@@

\textbf{Misalignments between stellar and dark matter distributions}
Figure \ref{fig:moneyplots} shows the differences between the angles of the stellar and the dark matter semi-major axes versus the ratio of the included dark matter and total mass. For lenses with a significant stellar component inside the outermost image radius, the distributions seem to be rather aligned. It seems that if the luminous mass is comparable to the dark matter mass, then the distributions align. For a smaller stellar contribution this is not necessarily the case. The scatter increases and there can be significant misalignments, which can be seen especially clearly in the systems \textit{PG1115} (Figure \ref{fig:1115rec}) and \textit{RXJ0911} (Figure \ref{fig:911}). While alignments between the distributions are preferred, they do not seem to be a necessary condition.

The environment might also play a role. Red are the systems, which are isolated or no group resp. cluster environment has been found, green the ones in such an environment. The misaligned systems are mainly the lenses with a known environment. Therefore, this misalignment could be caused by the external effects.

\textbf{Ellipticities of stellar and dark matter distributions}
The ellipticities of the distributions versus the mass ratio are shown in Figure \ref{fig:ellipticities}. We see, that every dark matter halo has a lower ellipticity $\lambda_{1}/\lambda_{2}$ and is thus closer to sphericity than the luminous galaxy. The systems \textit{HE1104} (Figure \ref{fig:1104}) and \textit{MG2016} (Figure \ref{fig:2016}) show especially large differences. It is not clear what could cause these

Figure \ref{fig:misalignmentellipticities} shows the misalignment versus the ellipticity of the dark matter halo. There seem to be five different regions. Almost spherical halos with good alignment of the semi-major axes, almost spherical halos with good alignment of semi-major and semi-minor axes (maximum misalignment between the semi-major axes), more elliptical, alligned halos, a seemingly 'forbidden' region for misalignments in between, and the upper right triangle, i.e. larger ellipticities and misalignments, which also seems forbidden. Understanding Figure \ref{fig:misalignmentellipticities} is certainly not trivial. Alignment with the semi-minor axis might be a stable configuration for almost spherical halos, while slightly smaller misalignments are unstable. There also seem to be some outliers as \textit{PG1115}, \textit{SBS1520}, and \textit{HE2149}.

\textbf{Misalignments between the dark matter distributions and environment}
Figure \ref{fig:environment} shows the misalignment between the direction to the group or cluster centroid resp. the companion galaxy with the dark matter halo. Although the error bars are huge, we immediately see that three lenses are basically aligned with the environment. The four other systems show slight misalignment. Having a closer look at these lenses shows however that at least two of them might in fact be aligned. For \textit{RXJ0911} and \textit{HE2149} there are no errors on the estimates of the centroid. Otherwise they would probably agree very well. \textit{Q0047} does not seem to be aligned with the direction of the group centroid. This could be due to the fact that the lens itself is quite far away from the centroid. The halo of \textit{B1030} is almost spherical, so the apparent misalignment could maybe be explained. It is also the only lens with only one companion galaxy. It is possible that in this case the effect of the environment is not strong enough. We however certainly lack data to make any claims for the systems with just one companion galaxy. Then, there is also \textit{MG2016}, which is not shown here, as the centroid was not measured in this cluster. There is however a high galaxy density towards the south-east, where also the cluster centroid probably lies. Judging from Figure \ref{fig:2016}, the halo seems to be aligned with the cluster centroid.

We conclude that lenses in groups or cluster are maybe aligned with the centroid. However, as we lack the statistics and as the errors on the centroid positions are huge, this claim might be disproved. It is not possible at this point to make a similar statement for lenses with just one companion galaxy.

>>>>

As mentioned in Section \ref{sec:data}, there are different time delay estimates for \textit{PG1115}. We find that the results do not change much when using different estimates. As can be seen in Figure \ref{fig:1115rec} resp. Figure \ref{fig:barkana}, the dark matter distribution is a bit more spherical with the time delays from \cite{2010MNRAS.406.2764T} and the uncertainty on the angle of the semi-major axis grows. The significant misalignment we see is thus not affected by different time delays.

@@

\textbf{Misalignments between stellar and dark matter distributions}
Figure \ref{fig:moneyplots} shows the differences between the angles of the stellar and the dark matter semi-major axes versus the ratio of the included dark matter and total mass. For lenses with a significant stellar component inside the outermost image radius, the distributions seem to be rather aligned. It seems that if the luminous mass is comparable to the dark matter mass, the distributions align. For a smaller stellar contribution this is not necessarily the case. The scatter increases and there can be significant misalignments, which can be seen especially clearly in the systems \textit{PG1115} (also Figure \ref{fig:1115rec}) and \textit{RXJ0911}. While alignments between the distributions are preferred, they do not seem to be the rule.

The environment might also play a role. Red are the systems, which are isolated and/or no group resp. cluster environment has been found yet, green the ones in such an environment. The misaligned systems are mainly the lenses with a known environment. Therefore, misalignments could be caused by the external effects.

\textbf{Ellipticities of stellar and dark matter distributions}
The ellipticities of the distributions versus the mass ratio are shown in Figure \ref{fig:ellipticities}. We see that every dark matter halo has a lower ellipticity $\lambda_{1}/\lambda_{2}$ and is thus closer to sphericity than the luminous part of the galaxy. The systems \textit{B0712}, \textit{B1152} \textit{HE1104}, and \textit{MG2016} show especially large differences. It is not clear what could cause these discrepancies.

Figure \ref{fig:misalignmentellipticities} shows the misalignment versus the ellipticity of the dark matter halo. It is possible to roughly subdivide it into five different regions. Almost spherical halos with good alignment of the semi-major axes, almost spherical halos with good alignment of semi-major and semi-minor axes (maximum misalignment between the semi-major axes), more elliptical, alligned halos, a seemingly 'forbidden' region for misalignments in between, and the upper right triangle, i.e. larger ellipticities and misalignments, which also seems forbidden. Understanding Figure \ref{fig:misalignmentellipticities} is certainly not trivial. Alignment with the semi-minor axis might be a stable configuration for almost spherical halos, while slightly smaller misalignments are unstable. The systems \textit{Q0957}, \textit{PG1115}, \textit{B1422}, \textit{MG2016}, and \textit{HE2149} should be studied more closely as they could hold valuable information to decipher these relations.

\textbf{Misalignments between the dark matter distributions and environment}
Figure \ref{fig:environment} we look at the alignment of the semi-major axes of the dark matter halos and vectors pointing to the centroid of the group or cluster or a companion galaxy. Although the error bars are huge, we see that five of these ten lenses, \textit{Q0047}, \textit{B0712}, \textit{Q0957}, \textit{PG1115}, and \textit{B1422}, are probably aligned or at least consistent with the environment. The other half show slight misalignment.

Having a closer look at these lenses shows however that at least two of them might in fact be aligned. For \textit{RXJ0911} and \textit{HE2149} there are no errors on the estimates of the centroid. Otherwise they would probably agree very well. \textit{Q0047} does not seem to be aligned with the direction of the group centroid. The halo of \textit{B1030} is almost spherical, so the apparent misalignment could maybe be explained. \textit{MG2016} also has a known environment, is however not shown in this Figure as no estimate of the cluster centroid is given. There is however a high galaxy density towards the south-east, where also the cluster centroid probably lies. The halo seems to be roughly aligned with the cluster centroid. %Add figure of this lens

We conclude that lenses in groups or cluster are probably aligned with the centroid, i.e. the center of mass, consistent with what one might expect naively. The light however needs not be as stated earlier.
%%%%%%%%%%%%%%%%%%
